Models, derivatives, response diversity

### 2024-07-08 models_derivatives_response_diversity.Rmd
re-examination of max derivatives

# Plot S3a: histogram of species' mean derivatives per lake
plot_percentiles <- df_quantiles |> 
  # Salmo trutta derivative is too large to be plotted
  filter(species != "Salmo_trutta") |> 
  ggplot(aes(x = mean_derivative, fill = factor(species))) + 
  geom_histogram(aes(y = after_stat(count / sum(count))), binwidth = 1.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("Cyprinus_carpio" = "#80CDC1",
                                "Alburnus_arborella" = "#8C510A",
                                "Barbatula_sp_Lineage_I"="#DFC27D",
                               "Phoxinus_csikii" = "#01665E",  "Salmo_trutta" = "#F46D43"))

@@

@@

>>>>

#  derivatives per lake
plot_percentiles <- df_quantiles |> 
  # Salmo trutta derivative is too large to be plotted
  filter(species != "Salmo_trutta") |> 
  ggplot(aes(x = mean_derivative, fill = factor(species))) + 
  geom_histogram(aes(y = after_stat(count / sum(count))), binwidth = 1.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("Cyprinus_carpio" = "#80CDC1",
                                "Alburnus_arborella" = "#8C510A",
                                "Barbatula_sp_Lineage_I"="#DFC27D",
                               "Phoxinus_csikii" = "#01665E",  "Salmo_trutta" = "#F46D43", "Cottus_sp_Po_profundal" = "red"))

@@

# LOOKING AT MAX DERIVATIVE
# Plot S3a: histogram of species' mean derivatives per lake
df_quantiles |> 
  # Salmo trutta derivative is too large to be plotted
  filter(species != "Salmo_trutta") |> 
  ggplot(aes(x = max_derivative, fill = factor(species))) + 
  geom_histogram(aes(y = after_stat(count / sum(count))), binwidth = 1.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("Cyprinus_carpio" = "#80CDC1",
                                "Alburnus_arborella" = "#8C510A",
                                "Barbatula_sp_Lineage_I"="#DFC27D",
                               "Phoxinus_csikii" = "#01665E",  "Salmo_trutta" = "#F46D43", "Cottus_sp_Po_profundal" = "red"))

@@

# max derivatives per species in each lake
max_deriv_plot <- data_quantiles |> 
  ggplot(aes(fLake, y = fct_reorder(species, max_derivative), fill= percentiles)) + 
  geom_tile() +
  scale_fill_manual(breaks = levels(data_quantiles$percentiles),
                    values = rev(brewer.pal(11, "BrBG")))

max_deriv_plot +
  guides(fill = guide_legend(title = "percentiles (max derivative)", reverse = TRUE)) +
  xlab("") +
  ylab("") +
  theme_classic(base_size = 16)
